Holomorphic field theories and higher algebra

### 2024-10-08 draft.tex
merge

\usepackage{parskip}
\setlength{\parindent}{18pt}
\setlength{\parindent}{0cm}

@@

\def\brian#1{{\textcolor{blue!75!red}{BW: {#1}}}}
\def\owen#1{{\textcolor{violet!50!black}{OG: {#1}}}}

@@

A community has formed, involving both mathematicians and physicists, to explore and exploit these relationships,
and it continues to grow as there is a lot to explore here.
Examples of efforts in this direction include \cite{PTVV}, \cite{FHK}, \cite{Keviny things}, \cite{GaiottoEtAl}, \cite{Paquette}, and \cite{BBBDN}, 
where our own background and interests (and unfortunately, our ignorance) have shaped this list.

The material below focuses on connections with this work,
and the examples from physics have a perturbative flavor.
We will not discuss important directions like mirror symmetry, the geometric Langlands program, the BPS/CFT correspondence, Donaldson-Thomas theory, or other areas that are more global or more directly geometric,
although we expect there are rich connections with the topics we do discuss.
Not including these topics is primarily because our core theme already swelled to more pages than we expected, 
because a concise discussion of any of those topics does not yet seem possible with respect to our theme,
and because of our lack of background in these areas.
We hope the ideas surveyed here do connect profitably to these topics.
\owen{Is that an overly negative paragraph?}

@@

For a systematic introduction to classical field theory, 
in a style amenable to a complex geometer,
we suggest \cite{DeligneFreed}.
There are many other sources, such as \cite{}, because classical field theory is an old and deep subject.
Recently there is an effort to revisit field theory with insights from derived geometry.
See \cite{AlfonsiYoung} for motivation and overview, 
but for a detailed development, see \cite{Steffens23, Steffens24}.
In Section~\ref{sec: pert dfn} we offer a sharp definition well-suited to perturbative field theory 
(i.e., studying formal deformations of a fixed solution to the equations of motion),
which we will use to produce examples of higher algebras.

\subsection{Examples}
\label{sec: ex of theories}

@@

There are holomorphic versions of all these basic types, 
and we now spell out some simple examples.\footnote{See \cite{DonThom} for a pioneering and illuminating discussion of how to find such analogs. See Nekrasov's thesis \cite{NekThesis} for a master physicist's view on holomorphic field theories, before he  reconfigured our understanding of much of gauge theory. \owen{What do you think of this?}}
To give the readers --- possibly a mixed audience of mathematicians and physicists --- some feel for the examples, 
we describe them in several ways: 
in terms of PDE and action functionals but also as moduli spaces.
We do not offer a thorough justification of the dictionary, for which see \cite{CosSUSY} for a wonderful and motivating exposition,
and see \cite{AlfonsiYoung} and \cite{Steffens23} for recent work on this topic.

@@

We will start with $M$ being a Riemann surface for simplicity.
Consider the action functional
\beqn
\label{eqn: 1d betagamma}
\int \beta \wedge \dbar \gamma
\eeqn
where $\beta$ is a $(1,0)$-form on $M$.\footnote{It can be viewed as a Lagrange multiplier for the field $\del \Bar{\gamma}$, 
which we now think about as having its own degrees of freedom. 
\owen{I don't understand this use of the term Lagrange multiplier. Let's discuss!}}
Note that this action functional also imposes the equation of motion $\dbar \beta = 0$,
so that $\beta$ should be a holomorphic $1$-form.

@@

Derived geometry is the mathematical version of BV formalism.
From this point of view, the moduli space of solutions for this $\beta\gamma$-system is the shifted cotangent bundle\owen{supervise notation!}
\[
T^*[-1] \Map_\dbar(M,\CC) = \RR\Gamma(M,\cO) \times \RR\Gamma(M,\cT^{*{1,0}}),
\]
a rather natural space to study in complex geometry.

@@

The $\beta\gamma$-system also plays a role in superstring theory as the free field theory underlying bosonic ghosts corresponding to worldsheet supersymmetry~\cite{Polchinski}.\owen{More precise location?}
There is likewise a complex-conjugate $\Bar{\beta}\Bar{\gamma}$-system encoding the antichiral sector of the free boson.
To recover the full boson, one must couple the chiral and anti-chiral sectors.
See \cite{Kapustin,GGW} \owen{Where?} for more details.~\hfill$\Diamond$
\end{rmk}

@@

From derived geometry point of view, the moduli space of solutions for this $\beta\gamma$-system is the shifted cotangent bundle\owen{supervise notation!}
\[
T^*[-1] \Map_\dbar(M,N) 
\]
a rather natural space to study in complex geometry.

@@

It is possible to formulate holomorphic versions of higher abelian gauge theories,
associated to moduli spaces of holomorphic $k$-gerbes with connection.
\owen{Should we elaborate or just stop with a reference?}

@@

%In complex dimension $3$ there is a related theory, called Kodaira--Spencer theory, which only consists of the field $\mu$ describing the complex structure in the above example.
Let $X$ be a complex threefold equipped with a Calabi--Yau structure; we denote by $\Omega$ the holomorphic volume form.
The primary field of Kodaira-Spencer theory is 
\beqn
\mu \in \Omega^{0,1}(X,\T_X),
\eeqn 
a $(0,1)$-form  on $X$ with values in the holomorphic tangent bundle~$\T_X$.
A constraint is imposed on these sections:
we assume that $\mu$ preserve the holomorphic volume form $\Omega$, meaning 
\[
{\rm div}_\Omega(\mu) = 0,
\]
i.e., its divergence vanishes.
In short, the fields are divergence-free vector fields.
The classical equation of motion is 
\beqn
\dbar \mu + \frac12 [\mu,\mu]_{SN} = 0 ,
\eeqn
where $[-,-]_{SN}$ denotes here the Schouten--Nijenhuis bracket.\footnote{On vector fields, $[-,-]_{SN}$ is simply the Lie bracket, and it is the natural multilinear extension to polyvector fields.}
The reader might recognize this equation as the Maurer--Cartan equation controlling deformations of complex structure $\dbar\rightsquigarrow \dbar + \mu$.
Formulating the action functional to obtain this equation of motion is a little involved, 
and we refer to \cite{CLbcov1} for a systematic discussion.

@@

This section offers a bridge from theories of physical interest to holomorphic field theories,
and hence may be of particular interest to physicists.

@@

There are many other appearances of holomorphic field theories in the context of supersymmetry.
Here is a (very) non-exhaustive list:
\begin{itemize}
\item Topological string theories yield rich classes of holomorphic field theories, as target space field theories. 
For instance, the topological B-model on a Calabi--Yau threefold $X$ leads to Kodaira--Spencer theory on $X$, as proposed in \cite{BCOV} using the framework of closed string field theory. 
The perspective of Kodaira--Spencer theory as a holomorphic field theory has been explored with great success by Kevin Costello and Si Li \cite{CLbcov, CLtypeI}.
\item
It is possible to make sense of Kodaira--Spencer theory outside of three dimensions.
In \cite{CLsugra}, Costello and Li have argued for conjectural descriptions of twists of supergravity and superstring theories in terms of this more general version of Kodaira--Spencer theory.
\item The pure spinor formalism of Berkovits and Cederwall  \cite{Cederwall, Berkovits} starts with a field theory or string theory defined on the (sometimes singular) nilpotence variety of the super Poincar\'e algebra and then uses the action of supersymmetry to produce ordinary (non-twisted) supersymmetric field theories. 
Often, the theory on the nilpotence variety is holomorphic \cite{ESW,SWpure}. 
\item Similar in spirit to the pure spinor formalism is a systematic relationship between holomorphic field theories on (super) twistor space and ordinary (Riemannian) field theories.
For instance, the Penrose corresponds relates holomorphic BF theory on (super) twistor space associated to $\RR^4$ to the self-dual limit of (super) Yang--Mills theory \cite{Penrose, Kevins work..., other twistor literature}.
\end{itemize}
We recommend \owen{is there some other place to point?} for more discussion.

@@

\owen{In this section, or maybe the next, we should mention the recent papers by physicists (notably Gaiotto) about higher bracket structures, and we should indicate how to connect with our approach here.}

@@

There are some features of this example that we would like to foreground. 
First, the commutative algebra arises as functions on a {\em symplectic} space,
which is the usual mathematical setting for classical mechanics (aside from more subtle situations that require Poisson geometry).
Second, the symplectic form $\omega = \d x \wedge \d p$ equips this algebra 
with a Poisson bracket where $\{ x, p \} = 1$,
which controls the deformation to the Weyl algebra: following Dirac, we promote the Poisson relation to a commutator relation.
These two features motivate the deformation quantization problem: 
given a symplectic (or Poisson) manifold,
describe deformations of its commutative algebra of functions to an associative algebra with the requirement that, to first order,  the commutator recovers the Poisson bracket.
Thanks to Kontsevich \cite{KonDQ}, there is a beautiful answer to this question, which has spawned a mountain of fascinating mathematics (see, e.g., \owen{what?} as a starting place).

Our view on field theory is motivated by this perspective on quantum theory (we discuss it further in Section~\owen{ref the general overview}), 
and it might help the reader to bear in mind a variant of this question:
given a classical holomorphic field theory, what are the natural deformations of its algebra of functions?
Many of the results we discuss later can be seen as generalizations of phenomena associated with deformation quantization.

@@

\brian{something about states}
So far we've recalled how quantization looks like for algebras.
There is also a quantization for modules.
Let $L \subset M$ be a Lagrangian submanifold.
Then, functions $C^\infty(L)$ on $L$ are naturally a module for $C^\infty(M)$.
The Fock module $F(L)$ associated to $L$ is defined to be the quotient of

\owen{We should acknowledge the state perspective but it's not clear to me how much we should describe, in light of our development below.}

@@

{\it Dimensional reduction}, an idea from physics, offers a view on how the Weyl algebra on generators $p, q$ relates to its loopy version.
As we already discussed, mechanics (the Weyl algebra of $V$) has to do with a one-dimensional system, i.e., time-line, 
while its loopy version ($W[V]$) has to do with a two-dimensional system, such as the punctured plane~$\CC^\times$.
There is a familiar identification (e.g., polar coordinates)
\[
\CC^\times \cong \RR \times S^1
\]
and so
\[
\Map(\CC^\times,V) \cong \Map(\RR, \Map(S^1,V)) = \Map(\RR, LV).
\]
Dimensional reduction here means studying the two-dimensional physics as a one-dimensional system with a more complicated target, namely the loop space.\footnote{One can play this game whenever the spacetime manifold is product $X \times Y$ or a fiber bundle.
Sometimes the term {\it compactification} is used instead, with reduction reserved for fixing the fiberwise fields to be constant.}
There is, in this case, a kind of dual process that we call {\it chiralization},
that turns a one-dimensional physical system (mechanics) into a holomorphic field theory.
We will now describe this relationship in the language of fields and Lagrangians.

@@

At the level of action functionals, 
we can integrate $S_{2d}(\phi)$ over the angular coordinate to obtain the following formal sum of one-dimensional action functionals
\beqn
S_{1d}\left( \ldots, \phi_{-1}, \phi_0 , \phi_1, \ldots \right) = \sum_{k\in \ZZ} \int  \phi_{-k} \,\d \phi_{k} + k\, \phi_k \, \phi_{-k} \, \d t .
\eeqn 
\owen{The $H$ got dropped. Should we reintroduct it?}
\brian{It's still there...}
Notice that the $k=0$ summand is just the Lagrangian of ordinary mechanics into $V$ with zero Hamiltonian. 
In general, we find mechanics into the symplectic vector space $LV$, where the symplectic structure uses the residue pairing in addition to the symplectic form on $V$, with a nontrivial Hamiltonian that encodes the $U(1)$-symmetry arising from rotating $LV$ in the natural way.

\owen{A person might object that this isn't so well-defined analytically. Should we emphasize that it is rather formal, a kind of heuristic?}

@@

There are natural central extensions of this sphere algebra 
%as {\em $L_\infty$ algebras},
parallel to the affine Lie algebras appearing as central extensions of the loop algebra.
They are given by the data of a $\fg$-invariant degree $d+1$ polynomial $\theta \in \Sym^{d+1}(\fg^*)^G$.
The cocycle that defines the $L_\infty$ central extension is
\[
\label{fhk cocycle}
\begin{array}{cccc}
\theta_{\rm FHK} : & (\sfA_d \tensor \fg)^{\tensor (d+1)} & \to & \CC\\ 
& a_0 \otimes \cdots \otimes a_d & \mapsto & \Res \left(\theta(a_0,\partial a_1,\ldots,\partial a_d)\right)
\end{array}.
\]
This cocycle has total cohomological degree $+2$ as an element in the Lie algebra cohomology of $\fg^\bu_{d}$ and so determines a central extension of dg Lie algebras
\[
\CC c \to \Hat{\fg}^\bullet_{d, \theta} \to \fg^\bullet_d .
\]
There is a particularly nice model for this dg Lie algebra as an $L_\infty$ \textit{algebra} that we use in \cite{GWkm}.
An $L_\infty$ algebra is a generalization of a dg Lie algebra;
it is a graded vector space $L$ equipped with a countable collection of ``higher brackets'' $[-]_k \colon L^{\otimes k} \to L$ that satisfy a system of equations generalizing the Jacobi identity.
For an overview (and an actual definition!) we recommend \owen{add citations}.
In our setting the usual bracket (the 2-bracket) of $\fg^\bullet_d$ is unchanged but there is now a nontrivial $d$-bracket defined by~$\theta_{\rm FHK}$.
%They dubbed these the {\em higher Kac-Moody algebras}, 
%as they specialize to the affine Lie algebras when~$d =1$.

@@

Associative algebras arise in quantum mechanics, we have argued, because the spacetime of a mechanical system is a line (``time'').\footnote{In the context of holomorphic field theory on $\C^n$, we have treated the radial direction as the ``time''.}
If we organize the observables (or operators) by when they happen in time,
it corresponds to writing them left-to-right as products in an algebra.
A natural question to ask then is what kind of higher algebraic structure captures how observables behave for theories in higher dimensions?
The shortest answer is ``factorization algebras,'' and we will describe and explore this notion in \S~\ref{sec fact alg},
but there are nice answers for two special classes of theories that we will develop now:
\begin{itemize}
\item For a topological field theory on $\RR^d$ (which we'll define in a moment),
the observables form an algebra over the {\em little $d$-discs operad}~$\Disc_d$ (sometimes also known as the $E_d$ operad). 
Equivalently, one says the observables are an $E_d$-algebra. 
This notion was introduced by algebraic topologists in the 1960s, 
so there is a rich body of work to draw upon.
\item For a holomorphic field theory on $\CC^d$,
the observables form an algebra over the $d$-polydiscs operad~$\PDisc_d$.
This notion was introduced only recently in~\cite{CG1}.
\end{itemize}
We will make the assertions above precise (in part by adding some natural but necessary hypotheses on a theory)
and develop them in parallel.
As a particularly illuminating example, we will explore the $d=1$ case rather closely and explain in what sense a vertex algebra is a kind of ``holomorphic associative algebra.''
We will also discuss the polydiscs algebras for which the higher Kac-Moody algebras are a Lie-algebraic shadow.

@@

There is an explicit model for an algebra over the cohomology of the $E_d$-operad called a $P_d$-algebra.
Essentially, these are graded commutative algebras with a Poisson bracket of cohomological degree $-d+1$.
In \cite{CostelloScheimbauer} a parallel model for the cohomology of the holomorphic $\PDisc_d$ colored operad was proposed, 
though axioms were not written down.
In recent work, similar models to that one have appeared,
using the nomenclature of $\lambda$-brackets \cite{BGKWY,GKW},
following Kac's approach in the $d=1$ setting. 
The multivariable $\lambda$-brackets are similar in spirit to Lie conformal algebras,
which use the notion of single-variable $\lambda$-bracket to encode the ``singular part'' of the data of a vertex algebra.
These papers use this algebraic structure to give a beautiful proposal for computing Feynman diagrams in holomorphic (and mixed holomorphic-topological) field theories (see Section~\ref{sec: renorm} for more on renormalization within the context of holomorphic QFT).

@@

We want to describe the algebraic structures that appear in $d$-dimensional topological field theories,
as a model for holomorphic field theories.
It will be helpful to use the language of operads, which are an efficient and powerful tool for defining and studying algebraic structures,
so we will gloss the key ideas of operads briefly and then turn to describing the operads relevant to TFTs.
A more extensive discussion can be found in \cite{CG1}, with \cite{ES} studying the topological setting in an elegant way.

@@

\begin{thm}[\cite{EllSaf}]
\label{thm: EllSaf}
Let $\cT$ be a classical field theory on $\RR^d$ in the sense of~\cite{CosBook},
and let its observables $\Obs_\cT$ be constructed following~\cite{CG2}.

@@

\owen{Improve!}
An algebra $A$ over this cooperad means that for each radius~$r$, 
there is a dg vector space $A(r)$, and for any list of radii $r_1,\ldots, r_n, R$,
there is a cochain map
\[
\alpha(r_1, \ldots, r_n| R): A(r_1) \otimes \cdots \otimes A(r_n) \to A(R) \otimes \Omega^{0,\bu}(\PDisc_d(r_1, \ldots, r_n| R)) .
\]
To invest the map $\alpha$ with more meaning, 
pick a real, closed submanifold 
\[
T \subset \PDisc_d(r_1, \ldots, r_n| R)
\]
along which one can integrate $(0,k)$-forms.
For instance, for any inputs $a_i \in A(r_i)$, we get an element in $A(R)$~by
\[
\int_T \alpha(a_1 \otimes \cdots \otimes a_n),
\]
where we have suppressed distracting notation from~$\alpha$.

@@

>>>>

\usepackage{parskip}
\setlength{\parindent}{18pt}
\setlength{\parindent}{0cm}
\setcounter{tocdepth}{1}

@@

\def\brian#1{{\textcolor{blue!75!red}{BW: {#1}}}}
\def\owen#1{{\textcolor{red!50!black}{OG: {#1}}}}

@@

A community has formed, involving both mathematicians and physicists, to explore and exploit these relationships,
and it continues to grow as there is a lot to explore here.
Examples of efforts in this direction include \cite{PTVV}, \cite{FHK,HenKap}, \cite{CosWG2,CosYangian,CosTwistor}, \cite{CLbcov, LiVertex}, \cite{CGholography,BGKWY}, \cite{CostelloPaquette}, and \cite{BBBDN}, 
where our own background and interests (and unfortunately, our ignorance) have shaped this list.
\brian{Maybe add more...}

The material below focuses on connections with this work,
and the examples from physics have a perturbative flavor.
We will not discuss important directions like mirror symmetry, the geometric Langlands program, the BPS/CFT correspondence, Donaldson-Thomas theory, or other areas that are more global or more directly geometric,
although we expect there are rich connections with the topics we do discuss.
Not including these topics is primarily because our core theme already swelled to more pages than we expected, 
because a concise discussion of any of those topics does not yet seem possible with respect to our theme,
and because of our lack of background in these areas.
We hope the ideas surveyed here do connect profitably to these topics.

@@

For a systematic introduction to classical field theory, 
in a style amenable to a complex geometer,
we suggest \cite{DeligneFreed}.
Recently there is an effort to revisit field theory with insights from derived geometry.
See \cite{AlfonsiYoung} for motivation and overview, 
but for a detailed development, see \cite{Steffens23, Steffens24}.
In Section~\ref{sec: pert dfn} we offer a sharp definition well-suited to perturbative field theory 
(i.e., studying formal deformations of a fixed solution to the equations of motion),
which we will use to produce examples of higher algebras.

\subsection{More examples}
\label{sec: ex of theories}

@@

There are holomorphic versions of all these basic types, 
and we now spell out some simple examples.\footnote{See \cite{DonThom} for a pioneering and illuminating discussion of how to find such analogs. See Nekrasov's thesis \cite{NekThesis} for a master physicist's view on holomorphic field theories, before he  reconfigured our understanding of much of gauge theory.}
To give the readers --- possibly a mixed audience of mathematicians and physicists --- some feel for the examples, 
we describe them in several ways: 
in terms of PDE and action functionals but also as moduli spaces.
We do not offer a thorough justification of the dictionary, for which see \cite{CosSUSY} for a wonderful and motivating exposition,
and see \cite{AlfonsiYoung} and \cite{Steffens23} for recent work on this topic.

@@

We will start with $M$ being a Riemann surface for simplicity.
Consider the action functional
\beqn
\label{eqn: 1d betagamma}
\int \beta \wedge \dbar \gamma
\eeqn
where $\beta$ is a $(1,0)$-form on $M$.%\footnote{It can be viewed as a Lagrange multiplier for the field $\del \Bar{\gamma}$, 
%which we now think about as having its own degrees of freedom. 
%\owen{I don't understand this use of the term Lagrange multiplier. Let's discuss!}}
Note that this action functional also imposes the equation of motion $\dbar \beta = 0$,
so that $\beta$ should be a holomorphic $1$-form.

@@

Derived geometry is the mathematical version of BV formalism.
From this point of view, the moduli space of solutions for this $\beta\gamma$-system is the shifted cotangent bundle
\[
T^*[-1] \Map_\dbar(M,\CC) = \RR\Gamma(M,\cO) \times \RR\Gamma(M,\Omega^1_{hol}),
\]
where $\Omega^1_{hol}$ is the sheaf of holomorphic one-forms.

@@

The $\beta\gamma$-system also plays a role in superstring theory as the free field theory underlying bosonic ghosts corresponding to worldsheet supersymmetry~\cite[Chapter 3]{Polchinski}.
There is likewise a complex-conjugate $\Bar{\beta}\Bar{\gamma}$-system encoding the antichiral sector of the free boson.
To recover the full boson, one must couple the chiral and anti-chiral sectors.
See \cite{Kapustin:2005pt,GGW} for more details.~\hfill$\Diamond$
\end{rmk}

@@

From derived geometry point of view, the moduli space of solutions for this $\beta\gamma$-system is the shifted cotangent bundle
\[
T^*[-1] \Map_\dbar(M,N) ,
\]
where $\Map_\dbar(M,N)$ is the derived stack of holomorphic maps $M \to N$.
This is a natural space to study in complex geometry.

@@

It is possible to formulate holomorphic versions of higher abelian gauge theories,
associated to moduli spaces of holomorphic $k$-gerbes with connection, see \cite{SWtensor,BWfive}.

@@

%In complex dimension $3$ there is a related theory, called Kodaira--Spencer theory, which only consists of the field $\mu$ describing the complex structure in the above example.
Let $X$ be a complex threefold equipped with a Calabi--Yau structure; we denote by $\Omega$ the holomorphic volume form.
The primary field of Kodaira-Spencer theory is 
\beqn
\mu \in \Omega^{0,1}(X,\T_X),
\eeqn 
a $(0,1)$-form  on $X$ with values in the holomorphic tangent bundle~$\T_X$.
A constraint is imposed on these sections:
we assume that $\mu$ preserve the holomorphic volume form $\Omega$, meaning 
\[
{\rm div}_\Omega(\mu) = 0,
\]
i.e., its divergence vanishes.
In short, the fields are divergence-free vector fields.
The classical equation of motion is 
\beqn
\dbar \mu + \frac12 [\mu,\mu]_{SN} = 0 ,
\eeqn
where $[-,-]_{SN}$ denotes here the Schouten--Nijenhuis bracket.\footnote{On vector fields, $[-,-]_{SN}$ is simply the Lie bracket, and it is the natural multilinear extension to polyvector fields.}
The reader might recognize this equation as the Maurer--Cartan equation controlling deformations of complex structure $\dbar\rightsquigarrow \dbar + \mu$.
Formulating the action functional to obtain this equation of motion is a little involved, 
and we refer to \cite{CLbcov} for a systematic discussion.

@@

This section offers a bridge from theories of physical interest to holomorphic field theories,
and hence may be of particular interest to physicists.
In his work on Donaldson invariants of four-manifolds, Witten introduced the idea of a topological twist of a supersymmetric field theory \cite{WittenTwist}.
The topological twist of a supersymmetric field theory (when it exists) can be attacked mathematically using the path integral.
A key feature, however, of topological field theories is that they are amenable to a functorial paradigm.
This firmly establishes their meaning and importance within pure mathematics.
Slightly after Witten's work, the first shadows of what would later be called \textit{holomorphic twists} of supersymmetric field theories appeared, see \cite{Johansen:1994aw,Johansen:1994ud,NekThesis,Losev:1996up,Losev:1995cr,Johansen:2003hw,Eager:2018oww}. 
Originally, this work did not garner the same attention as the parallel works on topological twists; especially in mathematics.
Much later, Costello introduced the precise notion of a holomorphic twist, and focused on examples in dimensions two and four \cite{CosSUSY}.
The characterization of all holomorphic, topological, and mixed holomorphic-topological theories of Yang--Mills type in dimensions $2 \leq d \leq 10$ appears in \cite{ESW}.
Holomorphic twists of more exotic quantum field theories, such as theories of supergravity and higher form gauge fields, have recently been studied in \cite{CLsugra,SWtensor,SWpure,Raghavendran:2021qbh,Eager:2021ufo,Garner:2023wrc}.

@@

There are many other appearances of holomorphic field theories in the context of supersymmetry.
Here is a (very) non-exhaustive list:
\begin{itemize}
\item Topological string theories yield rich classes of holomorphic field theories, as target space field theories. 
For instance, the topological B-model on a Calabi--Yau threefold $X$ leads to Kodaira--Spencer theory on $X$, as proposed in \cite{BCOV} using the framework of closed string field theory. 
The perspective of Kodaira--Spencer theory as a holomorphic field theory has been explored with great success by Kevin Costello and Si Li \cite{CLbcov, CLtypeI}.
\item
It is possible to make sense of Kodaira--Spencer theory outside of three complex dimensions.
In \cite{CLsugra}, Costello and Li have argued for conjectural descriptions of twists of supergravity and superstring theories in terms of this more general version of Kodaira--Spencer theory.
Holomorphic field theory has also appeared in applications to (twisted) holography through the lens of Koszul duality, see \cite{CLsugra, CGholography, CostelloPaquette}.
\item The pure spinor formalism of Berkovits and Cederwall  \cite{Cederwall:2001dx, Berkovits:2005bt}, see also \cite{Eager:2018dsx,perspectives}, starts with a field theory or string theory defined on the (sometimes singular) nilpotence variety of the super Poincar\'e algebra and then uses the action of supersymmetry to produce ordinary (non-twisted) supersymmetric field theories. 
Often, the theory on the nilpotence variety is holomorphic \cite{ESW,SWpure}. 
\item Similar in spirit to the pure spinor formalism is a systematic relationship between holomorphic field theories on (super) twistor space and ordinary (Riemannian) field theories.
Such work can be viewed as extending (and enhancing) the original Penrose correspondence which relates holomorphic BF theory on (super) twistor space associated to $\RR^4$ to the self-dual limit of (super) Yang--Mills theory \cite{Penrose:1969ae, Penrose}---see \cite{adamo2018lecturestwistortheory} for a modern review.
See \cite{Mason, Mason_2009, CostelloTwistor,Bittleston_2023,garner2024scatteringtwistoriallinedefects} for recent work on the twistor correspondence in QFT.
\end{itemize}

We recommend \cite{bah2024panoramaphysicalmathematicsc} for more discussion and references.

@@

@@

There are some features of this example that we would like to foreground. 
First, the commutative algebra arises as functions on a {\em symplectic} space,
which is the usual mathematical setting for classical mechanics (aside from more subtle situations that require Poisson geometry).
Second, the symplectic form $\omega = \d x \wedge \d p$ equips this algebra 
with a Poisson bracket where $\{ x, p \} = 1$,
which controls the deformation to the Weyl algebra: following Dirac, we promote the Poisson relation to a commutator relation.
These two features motivate the deformation quantization problem: 
given a symplectic (or Poisson) manifold,
describe deformations of its commutative algebra of functions to an associative algebra with the requirement that, to first order,  the commutator recovers the Poisson bracket.
Thanks to Kontsevich \cite{KonDQ}, there is a beautiful answer to this question, which has spawned a mountain of fascinating mathematics.

Our view on field theory is motivated by this perspective on quantum theory (we discuss it further in Section~\ref{s:factinqft}), 
and it might help the reader to bear in mind a variant of this question:
given a classical holomorphic field theory, what are the natural deformations of its algebra of functions?
Many of the results we discuss later can be seen as generalizations of phenomena associated with deformation quantization.

@@

%\brian{something about states}
%So far we've recalled how quantization looks like for algebras.
%There is also a quantization for modules.
%Let $L \subset M$ be a Lagrangian submanifold.
%Then, functions $C^\infty(L)$ on $L$ are naturally a module for $C^\infty(M)$.
%The Fock module $F(L)$ associated to $L$ is defined to be the quotient of 
%
%\owen{We should acknowledge the state perspective but it's not clear to me how much we should describe, in light of our development below.}

@@

{\it Dimensional reduction}, an idea from physics, offers a view on how the Weyl algebra on generators $p, q$ relates to its loopy version.
As we already discussed, mechanics (the Weyl algebra of $V$) has to do with a one-dimensional system, i.e., time-line, 
while its loopy version ($W[V]$) has to do with a two-dimensional system, such as the punctured plane~$\CC^\times$.
There is a familiar identification (e.g., polar coordinates)
\[
\CC^\times \cong \RR \times S^1
\]
and so
\[
\Map(\CC^\times,V) \cong \Map(\RR, \Map(S^1,V)) = \Map(\RR, LV).
\]
Dimensional reduction here means studying the two-dimensional physics as a one-dimensional system with a more complicated target, namely the loop space.\footnote{One can play this game whenever the spacetime manifold is product $X \times Y$ or a fiber bundle.
Sometimes the term {\it compactification} is used instead, with reduction reserved for fixing the fiberwise fields to be constant---for a discussion of compactification within the setting of factorization algebras see Section~\ref{s:compact}.}
There is, in this case, a kind of dual process that we call {\it chiralization},
that turns a one-dimensional physical system (mechanics) into a holomorphic field theory.
We will now describe this relationship in the language of fields and Lagrangians.

@@

At the level of action functionals, 
we can integrate $S_{2d}(\phi)$ over the angular coordinate to obtain the following formal sum of one-dimensional action functionals
\beqn
S_{red}\left( \ldots, \phi_{-1}, \phi_0 , \phi_1, \ldots \right) = \sum_{k\in \ZZ} \int  \phi_{-k} \,\d \phi_{k} + k\, \phi_k \, \phi_{-k} \, \d t .
\eeqn 
Notice that the $k=0$ summand is just the Lagrangian of ordinary mechanics into $V$ with zero Hamiltonian. 
In general, we find mechanics into the symplectic vector space $LV$, where the symplectic structure uses the residue pairing in addition to the symplectic form on $V$, with a nontrivial Hamiltonian that encodes the $U(1)$-symmetry arising from rotating $LV$ in the natural way.

%\owen{A person might object that this isn't so well-defined analytically. Should we emphasize that it is rather formal, a kind of heuristic?}

@@

There are natural central extensions of this sphere algebra 
%as {\em $L_\infty$ algebras},
parallel to the affine Lie algebras appearing as central extensions of the loop algebra.
They are given by the data of a $\fg$-invariant degree $d+1$ polynomial $\theta \in \Sym^{d+1}(\fg^*)^G$.
The cocycle that defines the $L_\infty$ central extension is
\[
\label{fhk cocycle}
\begin{array}{cccc}
\theta_{\rm FHK} : & (\sfA_d \tensor \fg)^{\tensor (d+1)} & \to & \CC\\ 
& a_0 \otimes \cdots \otimes a_d & \mapsto & \Res \left(\theta(a_0,\partial a_1,\ldots,\partial a_d)\right)
\end{array}.
\]
This cocycle has total cohomological degree $+2$ as an element in the Lie algebra cohomology of $\fg^\bu_{d}$ and so determines a central extension of dg Lie algebras
\[
\CC c \to \Hat{\fg}^\bullet_{d, \theta} \to \fg^\bullet_d .
\]
There is a particularly nice model for this dg Lie algebra as an $L_\infty$ \textit{algebra} that we use in \cite{GWkm}.
An $L_\infty$ algebra is a generalization of a dg Lie algebra;
it is a graded vector space $L$ equipped with a countable collection of ``higher brackets'' $[-]_k \colon L^{\otimes k} \to L$ that satisfy a system of equations generalizing the Jacobi identity.
For an overview (and an actual definition!) we recommend \cite{LV}.
In our setting the usual bracket (the 2-bracket) of $\fg^\bullet_d$ is unchanged but there is now a nontrivial $d$-bracket defined by~$\theta_{\rm FHK}$.
%They dubbed these the {\em higher Kac-Moody algebras}, 
%as they specialize to the affine Lie algebras when~$d =1$.

@@

Associative algebras arise in quantum mechanics, we have argued, because the spacetime of a mechanical system is a line (``time'').\footnote{In the context of holomorphic field theory on $\C^n$, we have treated the radial direction as the ``time''.}
If we organize the observables (or operators) by when they happen in time,
it corresponds to writing them left-to-right as products in an algebra.
A natural question to ask then is what kind of higher algebraic structure captures how observables behave for theories in higher dimensions?
The shortest answer is ``factorization algebras,'' and we will describe and explore this notion in Section~\ref{sec: fact alg}
but there are nice answers for two special classes of theories that we will develop now:
\begin{itemize}
\item For a topological field theory on $\RR^d$ (which we'll define in a moment),
the observables form an algebra over the {\em little $d$-discs operad}~$\Disc_d$ (sometimes also known as the $E_d$ operad). 
Equivalently, one says the observables are an $E_d$-algebra. 
This notion was introduced by algebraic topologists in the 1960s, 
so there is a rich body of work to draw upon.
\item For a holomorphic field theory on $\CC^d$,
the observables form an algebra over the $d$-polydiscs operad~$\PDisc_d$.
This notion was introduced only recently in~\cite{CG1}.
\end{itemize}
We will make the assertions above precise (in part by adding some natural but necessary hypotheses on a theory)
and develop them in parallel.
As a particularly illuminating example, we will explore the $d=1$ case rather closely and explain in what sense a vertex algebra is a kind of ``holomorphic associative algebra.''
We will also discuss the polydiscs algebras for which the higher Kac-Moody algebras are a Lie-algebraic shadow.

@@

There is an explicit model for an algebra over the cohomology of the $E_d$-operad called a $P_d$-algebra.
Essentially, these are graded commutative algebras with a Poisson bracket of cohomological degree $-d+1$.
In \cite{CosScheim} a parallel model for the cohomology of the holomorphic $\PDisc_d$ colored operad was proposed, 
though axioms were not written down.
In recent work, similar models to that one have appeared,
using the nomenclature of $\lambda$-brackets \cite{BGKWY,GKW},
following Kac's approach in the $d=1$ setting. 
The multivariable $\lambda$-brackets are similar in spirit to Lie conformal algebras,
which use the notion of single-variable $\lambda$-bracket to encode the ``singular part'' of the data of a vertex algebra.
These papers use this algebraic structure to give a beautiful proposal for computing Feynman diagrams in holomorphic (and mixed holomorphic-topological) field theories (see Section~\ref{sec: renorm} for more on renormalization within the context of holomorphic QFT).

@@

We want to describe the algebraic structures that appear in $d$-dimensional topological field theories,
as a model for holomorphic field theories.
It will be helpful to use the language of operads, which are an efficient and powerful tool for defining and studying algebraic structures,
so we will gloss the key ideas of operads briefly and then turn to describing the operads relevant to TFTs.
A more extensive discussion can be found in \cite{CG1}, with \cite{EllSaf} studying the topological setting in an elegant way.

@@

\begin{thm}[\cite{EllSaf}]
\label{thm: EllSaf}
Let $\cT$ be a classical field theory on $\RR^d$
and let its observables $\Obs_\cT$ be constructed following~\cite{CG2}.

@@

An algebra $A$ over this cooperad means that for each radius~$r$, 
there is a dg vector space $A(r)$, and for any list of radii $r_1,\ldots, r_n, R$,
there is a cochain map
\[
\alpha(r_1, \ldots, r_n| R): A(r_1) \otimes \cdots \otimes A(r_n) \to A(R) \otimes \Omega^{0,\bu}(\PDisc_d(r_1, \ldots, r_n| R)) .
\]
To invest the map $\alpha$ with more meaning, 
pick a real, closed submanifold 
\[
T \subset \PDisc_d(r_1, \ldots, r_n| R)
\]
along which one can integrate $(0,k)$-forms.
For instance, for any inputs $a_i \in A(r_i)$, we get an element in $A(R)$~by
\[
\int_T \alpha(a_1 \otimes \cdots \otimes a_n),
\]
where we have suppressed distracting notation from~$\alpha$.

@@

\subsection{A remark relating to homological projective duality}
